Querying Integrated Ecological Data

### 2010-04-29 docs/algdraft/DataIntegration/DataIntegration.tex
update more writing.

@@

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the conference you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals/conferences frown on
% math in the abstract anyway.

@@

\section{Introduction}
In this work, we study the query processing over integrated ecological
data through OBOE model\cite{DBLP:conf/er/BowersMS0}, a conceptual model used to interpret
observation and measurement data.

Contributions of this work:
\begin{itemize}
\item Formalize the queries over observation and measurement query.
\item Propose three methods to evaluate queries. 
\end{itemize}

\section{Related work}

This is related

\section{Data model and query}

In the ecological domain, observation and measurement data are the
core. As a consequence, people are more interested in finding datasets
related to specific observation and measurements.

@@

We a scientist contributes dataset into the data repository, generally
each dataset is converted to a data table. 
So, the database contains metadata about each dataset, the data table
definition (e.g., attribute/column name, attribute type, etc).

To formalize the senario, we use $D$ to denote the set of data tables. Let $d$ be one data table, 
$d[attr]$ or $d[idx]$ represents the set of values for the attribute
$attr$ or for the $idx$-th attribut.

Generally, the dataset is connected to the OBOE model using
annotations.

\from{HP}{Put the screendump of annotation.}

Map annotation information 
Through the annotation, 
We use $A$ to denote the annotation of one dataset.

\from{HP}{Another way to organize the data is to store them in one big
data table with three columns data\_file, attribute\_name,
attribute\_value. The problem with this organizationo is that the
attribute\_value are different types. }

%Implementation consideration: 
%When we put the data to data tables, we add a unique record id for
%each row in the data. This is used to get the unique record id.

@@

Given dataset in Figure \ref{}.

@@

We formalize the queries as follows.

\begin{definition}
KQ1: $<concept>:<cond>$
Here, <concept> can be observation, measurement, characteristic, standard,
unit, etc.. 
<cond> is in the form of $func([distinct]~attr)~op~value$, here $op$ can be some
algorithmmic operators $=$, $\neq$, $>$, or $<$. 
$func$ are aggregation functions including sum, average, min and max.

The result is a set of $d$, $\{d|d\in D \wedge (value~op~dv = true |dv\in d[attr])\}$.
\end{definition}

For each of such result, we denote that $d~s.t.~KQ$ meaning that
dataset $d$ satisfies the query.

This is the baseline keyword style query. 
The following two queries are the basic queries.

\begin{definition}
KQ2: $Q_1~context(Q_1.concept,Q_2)~Q_2 $, here $Q_i$ is a basic query
in the form of $<concept>:<cond>$. 
\end{definition}

\begin{example}
The given example queries can be represented in the formal query
format. 
\begin{itemize}
\item $Q_1 = Observation: Species='Picea rubens'$
\item $Q_2 = <Observation_1: Species='Picea rubens'>
  IN(Observation_1,Observation_2) <Observation_2:
  measurement_{unknown}=`California'>$
\item $Q_3 = Observation: Species='Picea rubens' \wedge
  count(distinct~Species)\geq 5$ \from{HP}{TODO}
\item $Q_4$ 
\end{itemize}

@@

@@

All the datasets can be imported to a RDBMS for

To answer such queries, a naive method is to

\subsection{Materialized database}

{\bf HP:}{Copy the materials from materiazeDB to here. Skip now.}

\subsection{Query rewriting}

@@

The first step is to find the dataset that contain all the metadata. 
From the annotation, find the {\em measurement type} with {\em
  Characteristic} species. 
Here, it is $m_3$. From the mapping, we find that the attribute in the
table is {\em spp}.

@@

%\bibliographystyle{abbrv}
%\bibliography{model}
%\bibliographystyle{IEEEtran}
%\bibliography{IEEEabrv,DataIntegration}

%\begin{thebibliography}{1}
%\bibliographystyle{abbrv}
%\bibliography{model}
%\end{thebibliography}

>>>>

\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

@@

\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}

@@

\section{Introduction}
In this work, we study the query processing over integrated ecological
data through OBOE model\cite{DBLP:conf/er/BowersMS08}, a conceptual model used to interpret
observation and measurement data.

\subsection{Background}
Introduce OBOE, annotation
Give an example of them.

\subsection{Contribution and paper organization}
Contributions of this work:
\begin{itemize}
\item Formalize the queries over observation and measurement query.
\item Propose three methods to evaluate queries. 
\end{itemize}

This paper is organized as follows. 
Section \ref{sec:relatedwork} reviews works that are related to this
research. Section \ref{sec:dataquery} formalize the data model and the
queries that people are interested to ask.

\section{Related work}\label{sec:relatedwork}

\section{Data model, annotation and query}\label{sec:dataquery}

In the ecology domain, most data people collected are observation and
measurement data. {\bf Add one sentence to explain this.}
Naturally, when searching ecological data, people are more interested in finding datasets
related to specific observation and measurement. 
In this section, we first illustrate the data model. Then, we
formalize the queries that scientists in this domain tend to ask.

@@

When a scientist contributes dataset into the data repository, generally
each dataset is converted to a data table\cite{tdar} or treated as a separate
object entity\cite{knb}. 
So, the database contains metadata about each dataset, the data table
definition (e.g., attribute/column name, attribute type, etc). 
%\from{HP}{Another way to organize the data is to store them in one big
%data table with three columns data\_file, attribute\_name,
%attribute\_value. The problem with this organization is that the
%attribute\_value are different types. } 
To formalize the scenario, we use $D$ to denote the set of data
tables. Each data table contains metadata about the attribute
definition.  Sometimes, we also use {\em column} to denote the attribute. 
Let $d$ be one data table, 
$d[attr]$ or $d[idx]$ represents the set of values for the attribute
$attr$ or for the $idx$-th attribute. 
Besides the dataset, in the data store, we also keep the annotation
information. 
%Implementation detail (no need to put it in the paper): 
%When we put the data to data tables, we add a unique record id for
%each row in the data. This is used to get the unique record id.

\subsection{Annotation}
In the Ecology domain, people are using annotations to connect
datasets to the OBOE model. 
{\bf TODO: Put a screen-dump of annotation.}

Map annotation information 
We use $A$ to denote the annotation of one dataset. 
More specifically,
To describe an observation, we need to denote on which entity (object in the
real world) the observation is made.
Very often, more than one observation can be made on one entity.   
\begin{itemize}
\item A.ObsType = $\{\langle \underline{AnnotId,ObsTypeId}, EntTypeId,
  isDistinct \rangle\}$ 
\item A.MeasType = $\{\langle \underline{AnnotId, MeasTypeId}, ObsTypeId,
  CharType, StdType, ProtType, Precision, isKey \rangle\}$ 
\item A.ContextType = $\{\langle \underline{AnnotId, ObsTypeId, ContextObsTypeId, RelType}, isIdentify\rangle\}$
\item A.Map = $\{\langle \underline{AnnotId,MeasTypeId, ResAttribute, Cond}, Val\rangle\}$
\end{itemize}

``AnnotId'' denotes the resource (e.g., data table) the annotation is on.

{\bf TODO:} Describe the constraints in annotation

@@

{\bf TODO: Have a running example from.}
Observation and measurement (OM) query

\begin{definition}[Basic OM query]\label{def:basic_omq}
A basic observation and measurement query is defined as $Q$::={\tt
  concept:cond}. 
Here, {\tt concept} is the main term in an observation model. 
In using the OBOE model, it can be observation, measurement,
characteristic or standard. 
{\tt cond} is in the form of $f([distinct]~attr)~op~value$ where $op$ is
basic operator from $\{=, \neq, >, <\}$ and $f$ is an
aggregation function from $\{sum, avg, count, min, max\}$. 
\end{definition}

\begin{definition}\label{def:qresult}
The result of a query $Q$ is a set of data objects (e.g., data tables) 
$\{d|d\in D \wedge d~satisfies~{\tt cond}\}$. 
%$\{d|d\in D \wedge (value~op~dv = true, dv\in d[attr])\}$. 
For each of such result $d$, we use $d~s.t.~Q$ to denote that
data object $d$ satisfies the query. 
\end{definition}

Definition \ref{def:basic_omq} can formulate queries on the
one specific observation and measurement. 
However, it fails to catch the observation context relationship. 
E.g., {\bf TODO: }
So, we generalize the basic OM query to contextualized OM query as
follows.

\begin{definition}[Contextualized OM query]\label{def:context_omq}
$CQ$::=$Q_1~${\tt context}($Q_1.concept, Q_2.concept$)$~Q_2$. Here $Q_i$ is a basic query
in the form of {\tt concept:cond}. 
\end{definition}

Given the dataset in Figure \ref{}. 
People may ask the following queries.

@@

They can be formalized as the following queries.

\begin{example}
The given example queries can be represented in the formal query
format. 
\begin{itemize}
\item $Observation: \langle Species='Picea~rubens'\rangle$
\item $Observation_1: \langle Species='Picea~rubens'\rangle$\\
  $IN(Observation_1, Observation_2$) \\
$Observation_2: \langle Meas_{unknown}=`California'\rangle$
\item $Observation: \langle Species='Picea~rubens' $\\
$\wedge
  count(distinct~Species)\geq 5\rangle$
\item $??$
\end{itemize}

@@

\section{Naive method}\label{sec:naive}
Index the data with all the values, then search over values. 
This may get wrong results.

{\bf Experiement: show the error rate (e.g., precision, recall, $F_1$ of this method.}

@@

As we analyzed in Section \ref{sec:relatedwork}, there are two extrems
in querying integrated data. 
One is to rewrite the original query to a series of new queries over
the data; the other is to materialize the data to a consistent data
model and then answer the query over the materialized database. 
In Section \ref{sec:matdb} and \ref{sec:queryrewrite}, we introduce
our strategies in answering the queries.

\subsection{Framework and operators}

\subsection{Querying over Materialized database}\label{sec:matdb}

\subsubsection{Materialize data}

\subsubsection{Querying materialized database}

\subsection{Query rewriting}\label{sec:queryrewrite}

For query rewriting, the first step is to find the candidate data
tables and the attribute name lists that can be used to answer our queries. 
To find the candidate data table, the metadata tables and annotation
tables are searched. 
When the candidate data tables and the needed attributes are known, 
the second step translate the needed queries over the candidate data
tables.

When a query {\tt concept:cond} is given, the {\tt concept} can only be interpreted using
annotations. When seeing a basic concept (observation, measurement,
characteristic, protocal), their related concepts information can be
retrieved through $A.MeasType$, $A.ObsType$.

From the {\em measuremen type} and $A.Map$, we can get the attribute
names; Then, using the attribute names and the metadata table (with
table and column definition), we can get the candidate data tables
with their corresponding attribute lists.

If there is context, we need to use $A.ContextType$.

If there is distinct constraint, need to find the key measurements. 
then, perform ``group by'' operation.

From the annotation, find the {\em measurement type} with {\em
  Characteristic} species. 
Here, it is $m_3$. From the mapping, we find that the attribute in the
table is {\em spp}.

@@

@@

\bibliographystyle{IEEEtranS}
\bibliography{IEEEabrv,DataIntegration}
